Enhancing Adversarial Robustness for Deep Metric Learning

### 2021-10-22 robdml.tex
tex

As suggested by previous works, an appropriate adversary for the inner
\emph{maximization} problem should increase the loss to a certain point between
that of benign examples (\ie, unperturbed examples) and the theoretical upper
bound.
%
What if this point is reached by the strongest adversary, directly and
efficiently?
%
Besides, as the mathematical expectation of loss is also greatly influenced by
the triplet sampling strategy, we conjecture that the triplet sampling strategy
and triplet ``hardness'' have a key impact for adversarial training.

@@

% HM
In this paper, we propose Hardness Manipulation (HM) to make adversarially
training a DML model with triplet loss feasible without model collapse, so the
model could efficiently learn from a strong adversary.
%
Assume there are two data triplets, namely ``source triplet'' and ``destination
triplet'' with the same anchor example, but sampled using different triplet
sampling strategies.
%
And the ``destination triplet'' is expected to result in a higher loss value
than another.
%
HM aims at creating adversarial examples of the ``source triplet'', that would
increase the corresponding loss value until that of the unperturbed
``destination triplet''.
%
This can also be interpreted as maximizing the loss value but ``stops early''
at a certain point.
%
The resulting perturbed ``source triplet'' can be used for adversarial
training of the model.

@@

Then, we want change (increase) the hardness level from $H_\mathsf{S}$ into $H_\mathsf{D}$, by
finding adversarial examples for $(X_a, X_p^S, X_n^S)$.
%
Note, it is not expected to lower the hardness during the attack, namely
$E[H_\mathsf{D}]$ should be greater or equal to $E[H_\mathsf{S}]$.
%
We denote the the hardness of adversarial examples as $\tilde{H}_S=H(X_a{+}r_a,
X_p^S{+}r_p, X_n^S{+}r_n)$.
%
This is named as \emph{Hardness Manipulation} (HM), which is implemented as
follows:
%
\begin{equation}
	%
	\bar{r}_a, \bar{r}_p, \bar{r}_n = \argmin_{r_a,r_p,r_n} \|\min(0,
	\tilde{H}_\mathsf{S}-H_\mathsf{D})\|^2.
	%
	\label{eq:hm}
	%
\end{equation}
%
The $\min(0,\cdot)$ part in \cref{eq:hm} truncates the gradient when $H_\mathsf{S}>H_\mathsf{D}$,
automatically stopping the optimization.
%
This is because $H_\mathsf{S}$ will be reduced once greater than $H_\mathsf{D}$ in standard Mean
Square Error.
%
The optimization problem can be solved by projected gradient
descent~\cite{madry}.
%
And the resulting adversarial examples are used for adversarially training the
DML model:
%
$L_\text{T}(\phi(X_a+\bar{r}_a), \phi(X_p^S+\bar{r}_p),
\phi(X_n^S+\bar{r}_n))$.

>>>>

As suggested by previous works, an appropriate adversary for the inner
\emph{maximization} problem should increase the loss to a certain point between
that of benign examples (\ie, unperturbed examples) and the theoretical upper
bound.
%
What if this point is reached by a strong adversary, directly and
efficiently?
%
Besides, as the mathematical expectation of loss is also greatly influenced by
the triplet sampling strategy, we conjecture that the triplet sampling strategy
and triplet ``hardness'' have a key impact for adversarial training.

% hardness manipulation

In this paper, we first define the ``\emph{hardness}'' of a sample triplet as
the difference between the anchor-positive distance and the anchor-negative
distance.
%
Then, Hardness Manipulation (HM) is proposed to adversarially perturb a given
sample triplet and increase its hardness into a specified hardness level.
%
HM is implemented as an objective to minimize the L-$2$ norm of the thresholded
difference between the hardness of the given sample triplet and the specified
hardness.
%
Mathematically, when the objective is optimized using Projected Gradient
Descent~\cite{madry}, the gradient of HM objective with respect to the
adversarial perturbation is the same as that of directly \emph{maximizing} the
triplet loss.
%
Thus, the optimization process of HM objective can be interpreted as a direct
\emph{maximization} process of the loss (indicating a direct strong adversary)
which stops halfway at the specified hardness level.
%
But, how hard should such ``specified level'' be?

by referencing
another benign (but harder) sample triplet.

make adversarially training a DML model
with triplet loss feasible without model collapse, so the model could
efficiently learn from a strong adversary.
%
Assume there are two data triplets, namely ``source triplet'' and ``destination
triplet'' with the same anchor example, but sampled using different triplet
sampling strategies.
%
And the ``destination triplet'' is expected to result in a higher loss value
than another.
%
HM aims at creating adversarial examples of the ``source triplet'', that would
increase the corresponding loss value until that of the unperturbed
``destination triplet''.
%
This can also be interpreted as maximizing the loss value but ``stops early''
at a certain point.
%
The resulting perturbed ``source triplet'' can be used for adversarial training
of the model.

@@

@@

Then, we want change (increase) the hardness level from $H_\mathsf{S}$ into $H_\mathsf{D}$, by
finding adversarial examples for $(X_a, X_p^S, X_n^S)$.
%
Note, it is not expected to lower the hardness during the attack, namely
$E[H_\mathsf{D}]$ should be greater or equal to $E[H_\mathsf{S}]$.
%
We denote the the hardness of adversarial examples as $\tilde{H}_S=H(X_a{+}r_a,
X_p^S{+}r_p, X_n^S{+}r_n)$.
%
This is named as \emph{Hardness Manipulation} (HM), which is implemented as
follows:
%
\begin{equation}
	%
	\bar{r}_a, \bar{r}_p, \bar{r}_n = \argmin_{r_a,r_p,r_n} \big\|\min(0,
	\tilde{H}_\mathsf{S} - H_\mathsf{D}) \big\|_2^2.
	%
	\label{eq:hm}
	%
\end{equation}
%
The $\min(0,\cdot)$ part in \cref{eq:hm} truncates the gradient when $H_\mathsf{S}>H_\mathsf{D}$,
automatically stopping the optimization.
%
This is because $H_\mathsf{S}$ will be reduced once greater than $H_\mathsf{D}$ in standard Mean
Square Error.
%
The optimization problem can be solved by projected gradient
descent~\cite{madry}.
%
And the resulting adversarial examples are used for adversarially training the
DML model:
%
$L_\text{T}(\phi(X_a+\bar{r}_a), \phi(X_p^S+\bar{r}_p),
\phi(X_n^S+\bar{r}_n))$.

\begin{align}
	&
	\sign \big\{
		-\frac{\partial}{\partial r} \big\| \min(0, 
		\tilde{H}_\mathsf{S} - H_\mathsf{D}) \big\|_2^2
	\big\}
	\\
	=&
	\sign \big\{
		-2(\tilde{H}_\mathsf{S}-H_\mathsf{D})
		\frac{\partial }{\partial r} \tilde{H}_\mathsf{S} \big\}
	=
	\sign \big\{
		\frac{\partial}{\partial r} \tilde{H}_\mathsf{S}
		\big\}
\end{align}
